Envelope Estimation using Geometric Properties of a Discrete Real Signal

### 2021-04-26 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
Arxiv version
reviewing

\documentclass{article}
\usepackage{lmodern} % Latin Modern Font
\usepackage[T1]{fontenc}
\usepackage{authblk}
\usepackage{amsfonts} % \mathbb
\usepackage[margin=0.5in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[backend=biber]{biblatex}
\usepackage{amsmath}

@@

\section{Introduction}
Envelope detection, also known as demodulation, is ubiquitous in both analogue and digital signal processing \parencite{2011CaetanoImproved}. Nevertheless, the literature in this area is fragmented \parencite{2017LyonsDigital}. Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise content, a limitation that excludes most physical signals, as is the case of recorded sound, for example; that limitation arises in part due to the lack of a strict mathematical definition of a temporal envelope \parencite{2013Mengempirical}.

@@

The envelope also helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions} and music: Concert halls with envelope preserving characteristics are deemed more pleasant \parencite{2011LokkiEngaging}, for example.

When dealing with broadband signals, approaches tailored to specific applications are prevalent, such as the one presented by \textcite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \textcite{2018AssefModeling} in the context of medical ultrasound imaging.

The problem of envelope detection, normally associated with the field of digital signal processing, can be made equivalent to the geometric problem of defining the (outer) shape of a set of points in $ \mathbb{R}^2 $, provided a method for addressing the difference of units in the horizontal, time related, and vertical, amplitude related, axes of a digital signal is available, rendering the original DSP problem readily approachable via geometric techniques.

One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the concept of a convex hull: as defined in \textcite{2008BergComputational}, the convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $. Convex hulls are unique, in the sense that a set of points define one and only one convex hull.

Generalizing that definition, \textcite{1983Edelsbrunnershape} introduced the concept of alpha shapes, that can be seen as a concave hull; a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.

The intuition behind the alpha shapes algorithm is that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the set of points of interest are part of the frontier, that is, of the concave hull of this set.

Once a radius $ \alpha $ is defined, alpha shapes conserve the uniqueness property, in fact tending to the convex hull when $ \alpha \to \infty $. Consider, for example, the artificially generated discrete wave in figure \ref{fig:Hulls} below, where the links between consecutive samples were deemphasized to highlight the “set of points” nature of a signal: The convex hull and a concave hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/00ConvexHull.pdf}
  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the convex and concave hull at $ \alpha=300 $ defined by it.}
  \label{fig:Hulls}
\end{figure}

@@

Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave, in order to completely avoid the need for manual intervention or parameter tuning.

@@

We follow with suggestions about how the approach here presented can be useful beyond envelope detection, exemplifying with some applications, such as the extractions of the average waveform of a digital wave, and concluding with commentary on potential future developments of the method.

\section{Methods}
After defining what will be considered an envelope in the context of this work, in this section we explore some simplifications that can be applied to a discrete wave, defining the representation that will be used in the subsequent part of the work.

\subsection{Characterization of a temporal envelope}

The interest in the theory of envelope detection arose in the analog domain, with the widespread adoption of radio communications, where the problem is typically restricted to simple sinusoids, often with well-defined frequencies \parencite{2011TurnerDemodulation}.

In this context, the most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, as introduced by Gabor in 1946 \parencite{2007HahnHistory}. In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
This analytic signal has the form $ A(t) = S(t) + \mathbb{H}(S(t)) \ \text{i} $ \parencite{2016HePraat} where $ S(t) $ is the original real signal. $ \mathbb{H}(S(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of the complex modulus of $ \mathbb{H}(S(t)) $.

@@

Figure \ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the shape of the envelope in the presence and absence of Gaussian white noise. It is readily noticeable from the figure that, as soon as noise is introduced in the original signal, it is reflected in the envelope.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/01AnalyticSignal.pdf}
  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach. In the first half, the sinusoid is free of noise, while in the second half Gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
  \label{fig:AnalyticSignal}
\end{figure}

@@

The preceding definitions can be translated to this discrete scenario assuming that the discrete quantities arise from observing the continuous ones at regular time intervals. In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each observation, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in the period of a second. We can thus define:

\begin{align} \label{eq:Envelope}
  & \mathbf{w}, \mathbf{e}, \mathbf{c} \in \mathbb{R}^n \\
  & \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) \\
  & \mathbf{e} = (e_0, e_1, \cdots, e_{n-1}) \\
  & \mathbf{c} = (c_0, c_1, \cdots, c_{n-1}) \\
  & \mathbf{w} = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
\end{align}

The $ \odot $ operator in \ref{eq:Envelope} stands for the Hadamard product, denoting element-wise multiplication of two vectors. Figure \ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by modulating the normalized carrier wave \textbf{c} using the envelope \textbf{e}.

Note from Figure \ref{fig:Envelope} that, at some local maxima, the envelope is well-defined, being in fact equal to the value of the wave $ w_i $, since $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.

The carrier wave \textbf{c} in the figure is periodic, and normalized between $ \{-1, 1\}$. Each cycle of the wave, then, reaches the value 1 only at its maximum making $ w_i $ equal to the value of the envelope at one point in each cycle.

Image \ref{fig:Envelope}, along with the concave hull illustrated in figure \ref{fig:Hulls}, helps forming the intuition behind the method developed in this work: the idea is to identify the local extrema that touch the envelope, those being also the extrema of each (pseudo-)cycle.

This is accomplished with the use of a circle with a defined radius, large enough to prevent it from reaching the smallest local extrema when "rolled" around the (geometric) representation of the wave. Only the extrema touched by the circle will be marked as part of the envelope.

In the next section, a compatible geometric representation will be formulated, along with the necessary tools to estimate the average curvature of the wave and, consequently, the appropriate radius of the circle.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/02Envelope.pdf}
  \caption{Example of a discrete wave \textbf{w} composed by the element-wise multiplication of an envelope \textbf{e} and a carrier \textbf{c}. The local extrema of \textbf{w} are highlighted with a circle.}
  \label{fig:Envelope}
\end{figure}

\subsection{Simplified Representation}

Instrumental to the method here presented is the definition of a pulse: each time the value of \textbf{w} changes sign, that is, every time the discrete wave \textbf{w} crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end. This definition is in line with the one presented in \textcite{national1996telecommunications}, where a pulse is defined as a rapid change in the amplitude of a signal, followed by a fast return to the baseline value; zero, in our case.

From \ref{fig:Envelope} and the discussion in the preceding chapter, we can see that only the maximum point of a positive pulse or the minimum point of a negative one can potentially be equal to the envelope and are, thus, our only points of interest. We can then proceed, for the rest of the method, considering only those points, to great computational economy.

@@

This relation is illustrated in figure \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/03Pulses.pdf}
  \caption{Example of a discrete wave \textbf{w} divided into pulses, of which the extrema are highlighted with a circle. P is the set of the points $ P_j = (i_j, \lvert w_j \lvert) $ representing the absolute value of those extrema.}
  \label{fig:Pulses}
\end{figure}

The points in P are not fit for geometric interpretation yet, since the abscissa and ordinate of their orthogonal coordinate system have different units: in the vertical axis, we have a unit related to the instantaneous intensity of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time-related unit.

Specifically, this coordinate system has a basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (a, 0), \mathbf{b}_y = (0, i) $ where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value. $ i $ represents the index of the sample, and is linked to time by the aforementioned relation $ i = t \ \text{fps} $.

We are interested in finding a coordinate system where a pulse's amplitude $ \lvert w_j \lvert $ and length would define, on average, a square.

@@

In this new coordinate system we have a basis vector $ \mathbf{b}^\prime = \{ \mathbf{b}^\prime_x, \mathbf{b}^\prime_y \}$ where $ \mathbf{b}^\prime_x = (i, 0) $ and $\mathbf{b}^\prime_y = (0, i) $, related to the original basis vector by equation \ref{eq:Basis}:

\begin{align} \label{eq:Basis}
\mathbf{b}^\prime_x &= \mathbf{b}_x  \\
\mathbf{b}^\prime_y &= \frac{\left( \frac{i_{m-1} - i_0}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)} \mathbf{b}_y 
\end{align}

The effect of this normalization can be seen in the points shown in figure \ref{fig:Vectors}, shown in scale in both coordinate systems. The set V of the vectors from each point of P to the next is also shown in the picture, and it will be used to estimate the average curvature of P. We define this set in equation \ref{eq:Vectors} below.

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/04Vectors.pdf}
  \caption{The set of points P and the set V of vectors between two adjacent points of P, in both the original and normalized coordinate systems, shown in scale.}
  \label{fig:Vectors}
\end{figure}

\subsection{Discrete Curvature Estimation}

Translating the explanation of alpha shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points that are part of the envelope are those touched by a circle of a given radius $ r $, coming from $ y \to \infty $ towards the signal, that is not allowed to contain any point of the signal; Intuitively, one can picture a circle being rolled above the signal, and marking the points it touches as envelope points.

To infer the appropriate radius $ r $ of such circle a measure of discrete curvature is needed. Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists.

The two possible general approaches are the derivation of direct methods, that use characteristics of the discrete wave to calculate the curvature, or the use of the curvature of a smooth, continuous curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}.

@@

Those approaches define the discrete curvature in relation to the vertices of a wave (or mesh, in the three-dimensional case). In our particular case, it wouldn't make sense to talk about the curvature of a single pulse, since it is what our points ultimately represent. Instead, we are interested in the change of direction, what the term curvature ultimately means, between two adjacent pulses, as the vectors in the set V suggests.

We thus proceed to define a discrete curvature measure over the edges of a discrete wave. To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve, noting that this is equivalent to that of the osculating circle \parencite{2016VasaMesh}.

\subsubsection{The Equivalent Circle Approach}

The rationale is to find, for each vector $ v_k \in \text{V} $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector of interest, as shown in figure \ref{fig:DiscreteCurvature}. The average curvature of P will then be obtained by the average of all radii $ r_k $.

From figure \ref{fig:DiscreteCurvature} is easy to see that $ r_k = v_{k,x} \sin(\theta_k)$ where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of a line in the horizontal direction and the slope $ m_k = v_{k,y} / v_{k,x} $ defined by $ v_k $ via the equality $ \theta_k = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/05DiscreteCurvature.pdf}
  \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta_k $ in $ \hat{u}_1 $, $ \theta_k $ being the angle that the vector $ v_k $ makes with the horizontal direction.}
  \label{fig:DiscreteCurvature}
\end{figure}

The radius of the circle that represents the average curvature of V (and P), and will be used to obtain the envelope of the discrete wave \textbf{w} can be obtained via equation \ref{eq:Radius}:

@@

We need now to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope. Generally, the algorithm for the alpha shapes approach resorts to the construction of the Delaunay triangulation of the set of points, that is latter filtered to contain only the outer edges. We can adopt a more straightforward method, as our points have a defined structure:

@@

For all circles of center $ x_o, y_o $ and radius $ r $ that can be constructed to pass through any two different points of P, that is, $ \forall \quad P_b, P_c $ such that $ (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2 $, $ P_a $ will be part of the set $ E $ of the points belonging to the envelope if and only if none of those circles contain $ P_a $, or $ P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2 $. Summarizing this, we have equation \ref{eq:InEnvelope}:

\begin{align} \label{eq:InEnvelope}
& P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2, \\
& \forall \ P_b, P_c \mid (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2, \\
& P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c \quad \text{and} \quad x_o, y_o \in \mathbb{R}, \quad 0 \le y_o < +\infty
\end{align}

@@

\section{Results and Discussion}
Given the lack of a precise definition of what an envelope should be, we begin this section suggesting an empirical metric, based on the behaviour of the carrier wave \textbf{c}. We also comment on some metrics presented in the literature, and compare our algorithm with traditional demodulation approaches, both in relation to execution time and accuracy; to this last end, a numerical indicator is also introduced.

@@

\subsection{Quality of an envelope}

There are no agreed-upon objective metrics to assess the quality of an envelope. Recalling figure \ref{fig:Envelope}, however, and the accompanying discussion, an intuitive way is to consider the carrier wave \textbf{c}, and how well it seems to be bounded by the interval $ \{-1, 1\} $.

@@

From figure \ref{fig:spoken_voice} it can be seen that the algorithm performs well under the carrier metric. This is especially remarkable for the case of spoken voice, a notorious complex wave with fast changes in pitch and high noise content.

These characteristics are in stark contrast with those of the sang voice in figure \ref{fig:singer_voice}, where a recording of an alto singer is shown; there, the wave is almost periodic, with a stable waveform.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_bend.pdf}
  \caption{Signal, envelope and carrier for a record of a bend performed on an electric guitar.}
  \label{fig:Bend}
\end{figure}

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_spoken_voice.pdf}
  \caption{Signal, envelope and carrier for a record of male voice uttering the word “amazing”.}
  \label{fig:spoken_voice}
\end{figure}

>>>>

\documentclass{article}
\usepackage{float}
\usepackage{lmodern} % Latin Modern Font
\usepackage[T1]{fontenc}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage{amsfonts} % \mathbb
\usepackage{amssymb} % \therefore
\usepackage[margin=0.5in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[backend=biber]{biblatex}

@@

% TODO only samples already present are part of the envelope; no overshooting
% TODO Frontiers are also provided

\section{Introduction}
Envelope detection, also known as demodulation, is ubiquitous in both analogue and digital signal processing (DSP) \parencite{2011CaetanoImproved}. Nevertheless, the literature in this area is fragmented \parencite{2017LyonsDigital}, with most envelope detection techniques being designed to account for very specific settings, like pure sinusoids with moderate noise content, for example.

Those constraints exclude most physical signals, as is the case of recorded sound, for example; this problem arises in part due to the lack of a strict mathematical definition of a temporal envelope \parencite{2013Mengempirical}.

@@

The temporal envelope also helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions} and music: Concert halls with envelope preserving characteristics are deemed more pleasant \parencite{2011LokkiEngaging}, for example.

Specially when dealing with broadband signals, approaches tailored to specific applications, exploiting a priori knowledge of the waves, are prevalent, such as the one presented by \textcite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \textcite{2018AssefModeling} in the context of medical ultrasound imaging.

The problem of envelope detection of a discrete signal, normally associated with the field of digital signal processing, can be made equivalent to the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.

Considering the standard 2d representation of a discrete wave, where the horizontal axis represents time and the vertical axis represents the amplitude, one can think of an individual sample as a point. Provided a coherent mapping from this space to the standard cartesian plane is available, the original DSP problem can be transformed into a geometric one.

% TODO

One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the concept of a convex hull, as defined in \textcite{2008BergComputational}. The convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $. Convex hulls are unique, in the sense that a given set of points define one and only one convex hull.

When the general shape of a set of points present prominent concavities, the convex hull overestimates the boundaries. To address this limitation \textcite{1983Edelsbrunnershape} introduced the concept of alpha shapes, that became known as the concave hull: a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.

The intuition behind the alpha shapes algorithm is that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the baricenter of the set of points are part of the frontier. The polygon obtained linking those points forms the concave hull of this set.

Once a radius $ \alpha $ is defined, alpha shapes exhibit the uniqueness property meaning that, for a given set of points and a given radius, only one concave hull exists. When the radius of the circle $ \alpha \to \infty $, concave hulls tend to the convex hulls.

Consider, for example, the artificially generated discrete wave in Fig. \ref{fig:Hulls} below, where the links between consecutive samples were deemphasized to highlight the "set of points" nature of a signal: The convex hull and a concave hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.

\begin{figure}[H]
  \centering
    \includegraphics{images/00ConvexHull.pdf}
  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the convex and concave hull at $ \alpha=300 $ defined by it.}
  \label{fig:Hulls}
\end{figure}

@@

Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a spectrally complex wave, in order to completely avoid the need for manual intervention or parameter tuning.

% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
After defining what will be considered an envelope in the context of this work, in this section we explore some simplifications that can be applied to a discrete wave, defining the representation that will be used in the subsequent part of the work.

@@

We follow with suggestions about how the approach here presented can be useful beyond envelope detection, exemplifying with some applications, such as the extractions of the average waveform of a digital wave, and concluding with commentary on potential future developments of the method.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Temporal envelope}\label{sec:Characterization of a temporal envelope}
The interest in the theory of envelope detection arose in the analog domain, with the widespread adoption of radio communications, where the problem is typically restricted to artificial waves, often with well-defined frequencies \parencite{2011TurnerDemodulation}.

% TODO vague definition of envelope - smoothness and slow varying

In this context, the most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, as introduced by Gabor in 1946 \parencite{2007HahnHistory}. In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.

This analytic signal has the form $ A(t) = S(t) + \mathbb{H}(S(t)) \ \text{i} $ \parencite{2016HePraat} where $ S(t) $ is the original real signal. $ \mathbb{H}(S(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of the complex modulus of $ \mathbb{H}(S(t)) $.

@@

Fig. \ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the smoothness of the envelope in the presence and absence of Gaussian white noise. It is readily noticeable from the Fig. \ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is directly reflected in the envelope.

\begin{figure}[H]
  \centering
    \includegraphics{images/01AnalyticSignal.pdf}
  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach. In the first half, the sinusoid is free of noise, while in the second half Gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
  \label{fig:AnalyticSignal}
\end{figure}

@@

The preceding definitions can be translated to this discrete scenario assuming that the discrete quantities arise from observing the continuous ones at regular time intervals. In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each observation, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in one second. We can thus define:

\begin{subequations}\label{eq:Envelope}
\begin{align}
  & \mathbf{w}, \mathbf{e}, \mathbf{c} \in \mathbb{R}^n \\
  & \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) \\
  & \mathbf{e} = (e_0, e_1, \cdots, e_{n-1}) \\  
  & \mathbf{c} = (c_0, c_1, \cdots, c_{n-1}) \\
  & \mathbf{w} = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
\end{align}
\end{subequations}

The $ \odot $ operator in \ref{eq:Envelope} stands for the Hadamard product, denoting element-wise multiplication of two vectors. Fig. \ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by modulating the normalized carrier wave \textbf{c} using the envelope \textbf{e}.

Note from Fig. \ref{fig:Envelope} that, at some local maxima, the envelope is well-defined, being in fact equal to the value of the wave $ w_i $. $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.

\begin{figure}[H]
  \centering
    \includegraphics{images/02Envelope.pdf}
  \caption{Example of a discrete wave \textbf{w} composed by the element-wise multiplication of an envelope \textbf{e} and a carrier \textbf{c}. The local extrema of \textbf{w} are highlighted with a circle.}
  \label{fig:Envelope}
\end{figure}

% TODO
This realization that the envelope of a wave is defined only at some of the wave's local extrema is crucial, as we can use it to generate a simplified representation of a signal that will make the algorithm more efficient by considering only the local extrema of a given signal.

\section{Simplified Representation}

We start with the definition of a pulse: each time the value of $ w_i$ changes sign, that is, every time the amplitude of the discrete wave \textbf{w} crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end. This definition is in line with the one presented in \textcite{national1996telecommunications}, where a pulse is defined as a rapid change in the amplitude of a signal, followed by a fast return to the baseline value; zero, in our case.

From \ref{fig:Envelope} and the discussion in Section \ref{sec:Characterization of a temporal envelope}, we can see that only the maximum point of a positive pulse or the minimum point of a negative pulse can potentially be equal to the envelope and are, thus, our only points of interest. We can then proceed, for the rest of the method, considering only those points, to great computational economy.

@@

This relation is illustrated in Fig. \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}.
While the exact relation between $ n $, the size of the original signal, and $ m $, the size of the set of points, is dependent on factors such as the frequency and sampling rate of the original wave, the inequality $ m \le n $ holds for all discrete signals. In practice, $ m \ll n $ for most real world discrete signals.

\begin{figure}[H]
  \centering
    \includegraphics{images/03Pulses.pdf}
  \caption{Example of a discrete wave \textbf{w} divided into pulses, of which the extrema are highlighted with a circle. P is the set of the points $ P_j = (i_j, \lvert w_j \lvert) $ representing the absolute value of those extrema.}
  \label{fig:Pulses}
\end{figure}

\section{Mapping to Cartesian coordinates}

The points in P are not fit for geometric interpretation yet, since the abscissa and ordinate of their orthogonal coordinate system have different units: in the vertical axis, we have a unit related to the instantaneous amplitude of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time-related unit.

Geometrically, this coordinate system has a basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (a, 0), \mathbf{b}_y = (0, i) $ where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value. $ i $ represents the index of the sample, and is linked to time by the aforementioned relation $ i = t \ \text{fps} $.

We are interested in finding a new basis vector that would make the geometric representation consistent, in the sense that pulse's amplitude $ \lvert w_j \lvert $, now mapped to the ordinates of a point, and a pulse's length, approximately equivalent to the horizontal distance between two consecutive points in the geometrical interpretation would define, on average, a square.

@@

In this new coordinate system we have a basis vector $ \mathbf{b}^\prime = \{ \mathbf{b}^\prime_x, \mathbf{b}^\prime_y \}$ where $ \mathbf{b}^\prime_x = (i, 0) $ and $\mathbf{b}^\prime_y = (0, i) $, related to the original basis vector by Eq. \ref{eq:Basis}:

\begin{subequations}  \label{eq:Basis} % TODO
\begin{align}
\mathbf{b}^\prime_x &= \mathbf{b}_x  \\
\mathbf{b}^\prime_y &= \frac{\left( \frac{i_{m-1} - i_0}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)} \mathbf{b}_y 
\end{align}
\end{subequations}

The effect of this normalization can be seen in the points shown in Fig.\ref{fig:Vectors}, shown in scale in both coordinate systems.

\section{Discrete Curvature Estimation}

We have now a map from the original space of the two-dimensional representation of a discrete wave to the Cartesian coordinate system and thus a way to consistently transform a discrete signal into a set of points in the plane.

Translating the explanation of the alpha shapes algorithm presented in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points that are part of the envelope are those touched by a circle of a given radius $ r $, coming from $ y \to \infty $ towards the Cartesian representation of the signal, that is not allowed to contain any point of the signal; Intuitively, one can picture a circle being rolled above the signal, and marking the points it touches as envelope points.

To employ the alpha shapes algorithm in this set of points, however, the radius $ r $ of such circle must be estimated, and for that a measure of discrete curvature is needed.

The set V of the vectors from each point of P to the next is also shown in the picture, and it will be used to estimate the average curvature of P. We define this set in Eq. \ref{eq:Vectors} below. % TODO

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/04Vectors.pdf}
  \caption{The set of points P and the set V of vectors between two adjacent points of P, in both the original and normalized coordinate systems, shown in scale.}
  \label{fig:Vectors}
\end{figure}

Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists.

The two general approaches are the derivation of direct methods that use characteristics of the discrete curve to calculate the curvature, or the use of the curvature of a smooth, continuous curve fitted to the original discrete curve \parencite{2001CoeurjollyDiscrete}.

@@

Those approaches define the discrete curvature in relation to the vertices of a discrete curve (or mesh, in the three-dimensional case). In our particular case, it wouldn't make sense to talk about the curvature of a single pulse, since it is what our points ultimately represent. Instead, we are interested in the change of direction, what the term curvature ultimately means, between two adjacent pulses, as the vectors in the set V suggests.

We thus proceed to define a discrete curvature measure over the edges of a discrete curve. To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve, noting that this is equivalent to that of the osculating circle \parencite{2016VasaMesh}.

\subsection{The Equivalent Circle Approach}

The rationale is to find, for each vector $ v_k \in \text{V} $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector of interest, as shown in Fig. \ref{fig:DiscreteCurvature}. The average curvature of P will then be obtained by the average of all radii $ r_k $.

From Fig. \ref{fig:DiscreteCurvature} is easy to see that $ r_k = v_{k,x} / \sin(\theta_k) $ where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of a line in the horizontal direction and the slope $ m_k = v_{k,y} / v_{k,x} $ defined by $ v_k $ via the equality $ \theta_k = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.

\begin{figure}[H]
  \centering
    \includegraphics{images/05DiscreteCurvature.pdf}
  \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta_k $ in $ \hat{u}_1 $, $ \theta_k $ being the angle that the vector $ v_k $ makes with the horizontal direction.}
  \label{fig:DiscreteCurvature}
\end{figure}

The radius of the circle that represents the average curvature of V (and, consequently, of P), and will be used to obtain the envelope of the discrete wave \textbf{w} can be obtained via Eq. \ref{eq:Radius}:

@@

We need now to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope. Generally, alpha shapes based algorithms resort to the construction of the Delaunay triangulation of the set of points, that is latter filtered to contain only the outer edges. We can adopt a more straightforward method, as our points have a sequential structure. Although that is the case throughout this work, the original signal doesn't need to be sampled in equal spaces.

@@

For all circles of center $ x_o, y_o $ and radius $ r $ that can be constructed to pass through any two different points of P, that is, $ \forall \quad P_b, P_c $ such that $ (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2 $, $ P_a $ will be part of the set $ E $ of the points belonging to the envelope if and only if none of those circles contain $ P_a $, or $ P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2 $. Summarizing this, we have Eq. \ref{eq:InEnvelope}:

\begin{subequations} \label{eq:InEnvelope}
\begin{align}
& P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2, \\
& \forall \ P_b, P_c \mid (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2, \\
& P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c \quad \text{and} \quad x_o, y_o \in \mathbb{R}, \quad 0 \le y_o < +\infty
\end{align}
\end{subequations}

@@

\section{Quality of an envelope}
Given the lack of a precise definition of what an envelope should be, we begin this section suggesting an empirical metric, based on the behaviour of the carrier wave \textbf{c}. We also comment on some metrics presented in the literature, and compare our algorithm with traditional demodulation approaches, both in relation to execution time and accuracy; to this last end, a numerical indicator is also introduced.

@@

There are no agreed-upon objective metrics to assess the quality of an envelope. Recalling Fig. \ref{fig:Envelope}, however, and the accompanying discussion, an intuitive way is to consider the carrier wave \textbf{c}, and how well it is bounded by the interval $ \{-1, 1\} $.

@@

From Fig. \ref{fig:spoken_voice} it can be seen that the algorithm performs well under the carrier metric. This is especially remarkable for the case of spoken voice, a notorious complex wave with fast changes in pitch and high noise content.

These characteristics are in stark contrast with those of the sang voice in Fig. \ref{fig:singer_voice}, where a recording of an alto singer is shown; there, the wave is almost periodic, with a stable waveform.

\begin{figure}[H]
  \centering
    \includegraphics{images/06Graph_spoken_voice.pdf}
  \caption{Signal, envelope and carrier for a record of male voice uttering the word “amazing”.}
  \label{fig:spoken_voice}
\end{figure}

\begin{figure}[H]
  \centering
    \includegraphics{images/06Graph_alto.pdf}
  \caption{Signal, envelope and carrier for a record of an alto singer.}
  \label{fig:singer_voice}
\end{figure}

Fig. \ref{fig:Bend} shows the envelope in the case of a bend performed in an eletric guitar. The variations in frequency in the signal didn't affect the envelope extracted.
